### 2018-01-29 MScFigures.Rmd
Added simplified analysis and data visualizations -- Rob

---
title: "MSc Statistical overview"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    highlight: tango
    theme: cerulean
    toc: yes
  word_document:
    toc: yes
---
# Housekeeping
### Preamble

By Katherine Duchesneau

We are working with the roots Lesion and Mycorrhiza dataset for now.

### R details

This section contains all the necessary details about the R version and package version used to run this Notebook.

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `tidyverse` package version: `r packageVersion("tidyverse")`
  * `lme4` package version `r packageVersion("lme4")`
  * `MASS` package version `r packageVersion("MASS")`
  * `MuMIn` package version `r packageVersion("MuMIn")`
  * `emmeans` package version `r packageVersion("emmeans")`
  * `boot` package version `r packageVersion("boot")`
  * `brms` package version `r packageVersion("brms")`
  * `loo` package version `r packageVersion("loo")`
  * `fitdistrplus` package version `r packageVersion("fitdistrplus")`
  * `mefa` package version `r packageVersion("mefa")`
  * `library(devtools)` and `install_github("vqv/ggbiplot")` package version `r packageVersion("ggbiplot")`
  * also requires [rtools](https://cran.r-project.org/bin/windows/Rtools/)

### Load Packages and Functions

```{r Load_packages and functions, message=FALSE, warning=FALSE}
library(tidyverse)
library(lme4)
library(MASS)
library(MuMIn)
library(emmeans)
library(boot)
library(brms)
library(loo)
library(fitdistrplus)
library(mefa)
library(ggbiplot)
```

## KAT, PLEASE BREAK UP YOUR CODE INTO SMALL CHUNCKS AND EXPLAIN WHAT EACH IS DOING. 
Load a custome theme for clean, readable graphs
```{r echo=F}
# theme_simple() a homemade function from the Colautti lab (2017) to standardize figures
theme_simple <- function (base_size = 16, base_family = "") {
  theme_classic(base_size = base_size, base_family = base_family) %+replace%
    theme(
      axis.text = element_text(colour = "black"),
      axis.title.x = element_text(size=18),
      axis.text.x = element_text(size=16, face= "italic"),
      axis.title.y = element_text(size=18, angle=90), 
      axis.text.y = element_text(size=16),
      axis.ticks = element_blank(),
      panel.background = element_rect(fill="white"),
      panel.border = element_blank(),
      plot.title=element_text(face="bold", size=24)
      #      legend.position="none"
    )
}
```

## EXPLAIN WHAT THIS OVERDISPERSION FUNCTION IS FOR...
```{r echo=F}
overdisp_fun <- function(model) {
  ## number of variance parameters in an n-by-n variance-covariance matrix
  vpars <- function(m) {
    nrow(m) * (nrow(m) + 1)/2
  }
  # The next two lines calculate the residual degrees of freedom
  model.df <- sum(sapply(VarCorr(model), vpars)) + length(fixef(model))
  rdf <- nrow(model.frame(model)) - model.df
  # extracts the Pearson residuals
  rp <- residuals(model, type = "pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  # Generates a p-value. If less than 0.05, the data are overdispersed.
  pval <- pchisq(Pearson.chisq, df = rdf, lower.tail = FALSE)
  c(chisq = Pearson.chisq, ratio = prat, rdf = rdf, p = pval)
}
```

# Introducing the Data

### Load/ Modify data: Co-occurence
```{r Introduce co-occurence dataset, message=FALSE, warning=FALSE}
mdata <- read.csv("Assessing_Oct_20.csv",stringsAsFactors=F)
indiv <- mdata$CODE[!mdata$CODE==""]
tdata <- mdata[grep("[0-9]+",mdata$Cross),]
nrow(tdata)/length(indiv)

# This is not the same because some individuals just had the totals not the cooccurence data
## THE COMMENT IS NOT CLEAR, EXPLAIN...
indiv2 <- tdata$CODE[!tdata$CODE==""]
nrow(tdata)/length(indiv2)
indiv2 <- data.frame(indiv2)
indiv2 <- mefa:::rep.data.frame(indiv2,each=100)
fdata <- cbind(indiv2,tdata)
fdata <- fdata[,-c(2,3,4,9,14,17,18)]

# Cleaning up the occurence data
## CLEANING UP HOW?
fdata[is.na(fdata)] <- 0
fdata$Lesion<-rowSums(fdata[,2:4])
fdata[fdata$Lesion>1,]
fdata$Mycorrhiza<-rowSums(fdata[,6:8])
fdata[fdata$Mycorrhiza>1,]
fdata <- fdata[,-c(2:4,6:8)]

#### Checking data for problems ####

summary(fdata)
fdata$None.Path <- as.integer(fdata$None.Path)
fdata$None.Myc <- as.integer(fdata$None.Myc)
fdata[is.na(fdata)] <- 0

sherb <- fdata$Herbivory+fdata$None.Herb
fdata[sherb!=1,]
fdata[sherb!=1,"None.Herb"]<-1

# Are all the sums equal to 100 for herb? 
dframeHerb <- data.frame(counter=NA, Indiv2=NA,Sum=NA)
c <- 1
for (i in unique(fdata$indiv2)){
  dframeHerb[c,] <- NA
  subdata <- fdata[fdata$indiv2==i,]
  dframeHerb$counter[c] <- c
  dframeHerb$Indiv2[c] <- i
  dframeHerb$Sum[c] <- sum(subdata[,"Herbivory"]+subdata[,"None.Herb"])
  subdata <- NA
  c <- c + 1
}
dframeHerb[dframeHerb$Sum!=100,]

# Are all the sums equal to 100 for Mycorrhiza? 
dframeMyc <- data.frame(counter=NA, Indiv2=NA,Sum=NA)
c <- 1
for (i in unique(fdata$indiv2)){
  dframeMyc[c,] <- NA
  subdata <- fdata[fdata$indiv2==i,]
  dframeMyc$counter[c] <- c
  dframeMyc$Indiv2[c] <- i
  dframeMyc$Sum[c] <- sum(subdata[,"Mycorrhiza"]+subdata[,"None.Myc"])
  subdata <- NA
  c <- c + 1
}
dframeMyc[dframeMyc$Sum!=100,]

# Are all the sums equal to 100 for Lesions? 
dframeLesion <- data.frame(counter=NA, Indiv2=NA,Sum=NA)
c <- 1
for (i in unique(fdata$indiv2)){
  dframeLesion[c,] <- NA
  subdata <- fdata[fdata$indiv2==i,]
  dframeLesion$counter[c] <- c
  dframeLesion$Indiv2[c] <- i
  dframeLesion$Sum[c] <- sum(subdata[,"Lesion"]+subdata[,"None.Path"])
  subdata <- NA
  c <- c + 1
}
dframeLesion[dframeLesion$Sum!=100,]

# Separating "Individual" into: pop,location,species columns 
fdata$pop <- gsub("^([0-9]+)[I,O][A-Z]+.*","\\1",fdata$indiv)
fdata$location <- gsub("^[0-9]+([I,O])[A-Z]+.*","\\1",fdata$indiv)
fdata$species <- sub("^[0-9]+[I,O]([A-Z]+).*","\\1",fdata$indiv)
fdata$location <- as.factor(fdata$location)
fdata$species <- as.factor(fdata$species)
fdata$pop <- as.factor(fdata$pop)
fdata$indiv2 <- as.character(fdata$indiv2)
summary(fdata)
fdata %>% drop_na()->fdata

fdata$None.Myc <- as.factor(fdata$None.Myc)
fdata$Herbivory <- as.factor(fdata$Herbivory)
fdata$Mycorrhiza <- as.factor(fdata$Mycorrhiza)
fdata$Lesion <- as.factor(fdata$Lesion)
fdata$None.Herb <- as.factor(fdata$None.Herb)
fdata$None.Path <- as.factor(fdata$None.Path)
summary(fdata)

#### Gather ####
length(unique(fdata$indiv2))
Cross<- data.frame(c(rep(1:100, times=225)))
fdatac <- cbind(fdata,Cross)
colnames(fdatac)[11]<-"cross"
fdatac %>% gather(Trait, Value, c(4,6,7)) -> fdata_gather # The warning is an inconsequential bug from tidyverse (https://github.com/tidyverse/tidyr/issues/386)
fdata_gather$Value<-as.numeric(fdata_gather$Value)
fdata_gather$Trait<-as.factor(fdata_gather$Trait)
fdata_gather <- fdata_gather[,-c(2:4)]
```

### Load/ Modify data: Totals
```{r Introduce total dataset, message=FALSE, warning=FALSE}
mdata<-read.csv("Assessing_Oct_20.csv",stringsAsFactors=F)
indiv<-mdata$CODE[!mdata$CODE==""]
tdata<-mdata[mdata$Slide=="TOTAL",]
length(indiv)
nrow(tdata)

ndata<-cbind(indiv,tdata)
ndata<-ndata[,-c(2,3,4,9,14,17:37)]
ndata$None.Path<-as.integer(ndata$None.Path)
ndata$None.Myc<-as.integer(ndata$None.Myc)

# Checking data for problems
problemsPath<-ndata[!rowSums(ndata[,c(2:5)])==100,]
problemsHyph<-ndata[!rowSums(ndata[,c(6:9)])==100,]
problemsHerb<-ndata[!rowSums(ndata[,c(10:11)])==100,]

# Separating "Individual" into: pop,location,species columns 
ndata$pop<-gsub("^([0-9]+)[I,O][A-Z]+.*","\\1",ndata$indiv)
ndata$location<-gsub("^[0-9]+([I,O])[A-Z]+.*","\\1",ndata$indiv)
ndata$species<-gsub("^[0-9]+[I,O]([A-Z]+).*","\\1",ndata$indiv)
ndata$location<-as.factor(ndata$location)
ndata$species<-as.factor(ndata$species)
ndata$pop<-as.factor(ndata$pop)

#### Totalling the scores####
binlesion<-cbind(ndata$None.Path,100-ndata$None.Path)
binmycorr<-cbind(ndata$None.Myc,100-ndata$None.Myc)
binherbivory<-cbind(ndata$None.Herb,100-ndata$None.Herb)
summary(ndata)

# scale data (only need to scale the numerical variables!)
numcols <- grep("^None.",names(ndata))
ndatas <- ndata
ndatas[,numcols] <- scale(ndatas[,numcols])
```

### Variable descriptions

The variables are:

Co-occurence:

  * **indiv2:** The unique code given to a sample. In this dataset the unique sample code repeats 100 times for each cross where an observation was taken on the root sample.

  * **None.Path:** A binary representation of whether a sign of pathogen activity was recorded (0) or not (1).

  * **Lesion:** The counterpart to the previous variable (None.Path). A binary representation of whether a sign of pathogen activity was recorded (1) or not (0).

  * **None.Myc:** A binary representation of whether a sign of myccorhizal activity was recorded (0) or not (1).

  * **Mycorrhiza:** The counterpart to the previous variable (None.Myc). A binary representation of whether a sign of mycorrhizal activity was recorded (1) or not (0).

  * **None.Herb:** A binary representation of whether herbivory was recorded (0) or not (1).

  * **Herbivory:** The counterpart to the previous variable (None.Herb). A binary representation of whether Herbivoryn was recorded (1) or not (0).

  * **Population (pop):** The coding number representing the population at which the sample was collected.

  * **location:** A code representing the whether the sample was collected inside a Alliaria petiolata population (I) or whther it was collected at least 7 m outside of he furthest individual in the A. petiolata population (O).

  * **species:** The particular species to which the sample belongs.

  * **Cross:** The particular cross number where the data was recorded on the individual sample.

Total scores:

  * **indiv:** The unique code given to a sample.

  * **Decay:** The total number of crosses where decay was recorded an individual sample when doing a total of 100 crosses/ sample.

  * **Pathogen:** The total number of crosses where pathogen was recorded on an individual sample when doing a total of 100 crosses/ sample.

  * **Hyphae:** The total number of crosses where non-myccorhizal hyphae was recorded on an individual sample when doing a total of 100 crosses/ sample.

  * **None.Path:** The total number of crosses where no signs of pathogen activities were recorded on an individual sample when doing a total of 100 crosses/ sample. Note that the total of Decay, Pathogen, Hyphae, and None.Path must come to 100 per individual to account for all 100 crosses.

  * **Arbuscule:** The total number of crosses where an arbuscule was recorded on an individual sample when doing a total of 100 crosses/ sample.

  * **Vesicules:** The total number of crosses where a vesicule was recorded on an individual sample when doing a total of 100 crosses/ sample.

  * **M_Hyphae:** The total number of crosses where myccorhizal hyphae was recorded on an individual sample when doing a total of 100 crosses/ sample.

  * **None.Myc:** The total number of crosses where no signs of mycorrhizal activities were recorded on an individual sample when doing a total of 100 crosses/ sample. Note that the total of Arbuscules, Vesicules, M_Hyphae, and None.Myc must come to 100 per individual to account for all 100 crosses.

  * **Herbivory:** The total number of crosses where herbivory was recorded on an individual sample when doing a total of 100 crosses/ sample.

  * **None.Herb:** The total number of crosses where no signs of herbivory was recorded on an individual sample when doing a total of 100 crosses/ sample. Note that the total of Herbivory, and None.Herb must come to 100 per individual to account for all 100 crosses.

  * **Population (pop):** The coding number representing the population at which the sample was collected.

  * **location:** A code representing the whether the sample was collected inside a Alliaria petiolata population (I) or whther it was collected at least 7 m outside of he furthest individual in the A. petiolata population (O).

  * **species:** The particular species to which the sample belongs.

# Reviewing of hypotheses:

## Question 1: Is there evidence that Mycorrhizal colonization provides protection against root pathogen damage?

## Question 2: Do these effects vary among species?

### Hypothesis 1

### VERY NICE PREDICTION FIGURES! PERSONALLY, I WOULD MAKE THE DOTS BIGGER AND BIVARIATE NORMAL (SEE CHANGES):

Mycorrhizal colonization protects roots from lesions and herbivory.
```{r echo=FALSE}
set.seed(3242)
y<-data.frame(rnorm(100,sd=7))
colnames(y)<-c("y")
y$y<-sort(y$y)
x<-data.frame(-y+rnorm(100))
colnames(x)<-c("x")
df <- cbind(x,y)
colnames(df)<-c("x","y")
df$Location <- c(rep("Out",nrow(df)/2),rep("In",nrow(df)/2))

ggplot(data=df, aes(x,y))+geom_smooth(method="lm", se=F)+geom_point(size=4,alpha=0.3)+
  theme_simple()+labs(x="Mycorrhizal Colonization (%)", y="Lesions (%)")+
  theme(axis.text.x = element_blank(),axis.text.y = element_blank())
```

### Hypothesis 2

A. petiolata invasion reduces mycorrhizal colonization and increases pathogen colonization and herbivory
```{r echo=FALSE}
ggplot(df, aes(x,y, colour=Location))+geom_smooth(method="lm", se=F)+
  geom_point(size=4,alpha=0.3)+theme_simple()+
  labs(x="Mycorrhizal Colonization (%)", y="Lesions (%)")+
  theme(axis.text.x = element_blank(),axis.text.y = element_blank())
```
### Hypothesis 3

Cycorrhizal colonization is reduced in the presence of A. petiolata
```{r echo=FALSE}
df$y2<-df$y
df$y2[df$Location=="In"]<-df$y2[df$Location=="In"]+10
ggplot(df, aes(x,y2, colour=Location))+geom_smooth(method="lm", se=F)+
  geom_point(size=4,alpha=0.3)+
  theme_simple()+labs(x="Mycorrhizal Colonization (%)", y="Lesions (%)")+
  theme(axis.text.x = element_blank(),axis.text.y = element_blank())
```

### Hypothesis 4

Mycorrhizal protection is completely lost in the presence of A. petiolata
```{r echo=FALSE}
df$y2<-df$y
set.seed (238.4)
df$y2[df$Location=="In"]<-rnorm(nrow(df)/2,mean=10)
ggplot(df, aes(x,y2, colour=Location))+geom_smooth(method="lm", se=F)+
  geom_point(size=4,alpha=0.3)+
  theme_simple()+labs(x="Mycorrhizal Colonization (%)", y="Lesions (%)")+
  theme(axis.text.x = element_blank(),axis.text.y = element_blank())
```

## Molecular data Hypothesis

### Hypothesis 1

A. petiolata invasion changes the diversity of mycorrhizal and other fungal species present in the soil.

I will find that the molecular data from the soil extraction supports this claim by having a reduced biodiversity and abundance of mycorrhiza in samples taken inside A. petiolata populations.

## IMPORTANT: BEFORE RUNNING COMPLICATED MODELS: (1) TAKE A LOOK AT THE DATA, (2) RUN SIMPLE MODELS -- SEE IF IT IS EVEN WORTH YOUR TIME & EFFORT

### Hypothesis 2

There will be a negative correlation between mycorrhizal and fungal pathogen diversity.

# Histology data Analysis

### Data exploration:

Calculate plant-level Mycchorizae, pathogens and herbivory:

```{r}
# Fix column characteristics for plotting and analysis
fdata$indiv2<-as.factor(fdata$indiv2)
fdata$Mycorrhiza<-as.numeric(paste(fdata$Mycorrhiza))
fdata$Lesion<-as.numeric(paste(fdata$Lesion))
fdata$Herbivory<-as.numeric(paste(fdata$Herbivory))
pct<-function(x){
  sum(x)/100
}

# Calculate plant-level averages
fdat_ind<-aggregate(fdata[,c("Mycorrhiza","Lesion","Herbivory")],
                    by=list(indiv2=fdata$indiv2,pop=fdata$pop,loc=fdata$loc,species=fdata$species),
                    pct)

```

## SIMPLE LINEAR MODEL AND VISUALIZATION:

### Mycorrhia model results

  1. Mycorrhiza colonization is significantly higher outside patches of A. petiolata than inside
  2a. Species vary in Mycorrhiza colonization, with highest colonization in SC and others relatively similar. 
  2b. Effect of A. petiolata on Mycorrhiza colonization differs among species, with more pronounced effects for GR and MR, and will little effect for GA
  3. Effect of A. petiolata on Mycorrhiza colonization varies by population, but not as strongly as among species

```{r}
## Mycorrhiza model (NOTE: No loc*species because not all species present in all populations)
base.Myc<-lmer(Mycorrhiza~species+loc+species:loc+(1|pop)+(1|pop:loc),data=fdat_ind)
noint.Myc<-lmer(Mycorrhiza~species+loc+(1|pop)+(1|pop:loc),data=fdat_ind)
anova(base.Myc,noint.Myc)
base.Myc ## Inspect coefficients

ggplot(fdat_ind,aes(species,Mycorrhiza,colour=loc))+geom_boxplot()+theme_simple()
ggplot(fdat_ind,aes(pop,Mycorrhiza,colour=loc))+geom_boxplot()+theme_simple()
```

### Lesions model results

  1. Root lesions increase with mycorrhiza colonization, contrary to predictions
  2. Effect of Mycorrhiza on Lesions depends strongly on species (Myc*species interactions)

```{r}
## Mycorrhiza model (NOTE: No loc*species because not all species present in all populations)
base.Les<-lmer(Lesion~Mycorrhiza*species*loc+(1|pop/loc),data=fdat_ind)
norint.Les<-lmer(Lesion~Mycorrhiza*species*loc+(1|pop),data=fdat_ind)
anova(base.Les,norint.Les)
norint.Les # inspect coefficients

Les2<-lmer(Lesion~Mycorrhiza*species*loc-Mycorrhiza:species:loc+(1|pop),data=fdat_ind)
anova(norint.Les,Les2)

Les3a<-lmer(Lesion~Mycorrhiza*species+Mycorrhiza*loc+(1|pop),data=fdat_ind)
Les3b<-lmer(Lesion~Mycorrhiza*species+species*loc+(1|pop),data=fdat_ind)
Les3c<-lmer(Lesion~species*loc+Mycorrhiza*loc+(1|pop),data=fdat_ind)
anova(norint.Les,Les3a) # TEST species*loc 
anova(norint.Les,Les3b) # TEST Myc*loc 
anova(norint.Les,Les3c) # TEST Myc*species

ggplot(fdat_ind,aes(Mycorrhiza,Lesion))+facet_wrap(~species)+
  geom_point(aes(col=loc))+theme_simple()+geom_smooth(method="lm")
```

## Hypothesis 1: Lesion ~ Mycorriza

### GIVEN THE NEW ANALYSIS ABOVE, THIS ISN'T WORTHWILE BECAUSE THERE IS NO SIGNIFICANT Myc:species:loc

```{r}
# calculating the r
indivCODE <- unique(fdata$indiv2)
dfcorr <-data.frame(indivCODE=c(0),r=c(0))
for (i in 1:length(indivCODE)){
  dat1<-NULL
  dat1<-filter(fdata, fdata$indiv2==indivCODE[i])
  dfcorr[i,1] <- indivCODE[i]
  dfcorr[i,2] <- cor(as.numeric(dat1$Mycorrhiza),as.numeric(dat1$Lesion))
}
dfcorr %>% drop_na()->dfcorr
```

###Co-occurence data

```{r Co-occurence model using corr}
#Normal?
car::qqp(dfcorr$r, "norm")
shapiro.test(dfcorr$r)
hist(dfcorr$r)
# Cannot be normal because it is only between -1 and 1 and our mean is at 0.
descdist(dfcorr$r, discrete = FALSE, boot = 500)

# weibull when transformed to be positive?
dfcorr$r.t <- dfcorr$r+1
fit.weibull <- fitdist(dfcorr$r.t, "weibull")
car::qqp(dfcorr$r.t, "weibull", shape = fit.weibull$estimate[[1]])

dfcorr$pop <- gsub("^([0-9]+)[I,O][A-Z]+.*","\\1",dfcorr$indivCODE)
dfcorr$location <- gsub("^[0-9]+([I,O])[A-Z]+.*","\\1",dfcorr$indivCODE)
dfcorr$species <- sub("^[0-9]+[I,O]([A-Z]+).*","\\1",dfcorr$indivCODE)
dfcorr$location <- as.factor(dfcorr$location)
dfcorr$species <- as.factor(dfcorr$species)
dfcorr$pop <- as.factor(dfcorr$pop)

meanR<-mean(dfcorr$r)
SEMR<-sd(dfcorr$r)/sqrt(length(dfcorr$r))
UCI<- meanR+SEMR*1.96
LCI<- meanR-SEMR*1.96
mean(filter(dfcorr, dfcorr$location=="I")$r)
mean(filter(dfcorr, dfcorr$location=="O")$r)

baysian_corrGLMM <- brm(r.t ~ location * species + (1|pop), data=dfcorr, family = weibull(link="log"), control = list(adapt_delta = 0.9))
baysian_corrGLMM2 <- brm(r.t ~ location + species + (1|pop), data=dfcorr, family = weibull(link="log"), control = list(adapt_delta = 0.9))
baysian_corrGLMM3 <- brm(r.t ~ location + (1|pop), data=dfcorr, family = weibull(link="log"), control = list(adapt_delta = 0.9))
baysian_corrGLMM_NULL <- brm(r.t ~ 1 + (1|pop), data=dfcorr, family = weibull(link="log"), control = list(adapt_delta = 0.9))

waic.BCG_1<-waic(baysian_corrGLMM)
waic.BCG_2<-waic(baysian_corrGLMM2)
waic.BCG_3<-waic(baysian_corrGLMM3)
waic.BCG_NULL<-waic(baysian_corrGLMM_NULL)
compare(waic.BCG_1,waic.BCG_2) #positive means second is better
compare(waic.BCG_2,waic.BCG_3) #negative means first is better
compare(waic.BCG_2,waic.BCG_NULL) #negative means first is better
summary(baysian_corrGLMM2)
bayes_R2(baysian_corrGLMM2)

ggplot(dfcorr, aes(y=r,x=location))+geom_boxplot()+theme_simple()
ggplot(dfcorr, aes(y=r,x=species))+geom_boxplot()+theme_simple()

```

Co-occurence with gather: what is the difference? 
Here I want to see that the traits are associating differently at every cross for each indiv

````{r Co-occurence model GATHER}

baysianGLMERmyc_lesion_GTHR <- brm(Value ~ Trait * species + (1|pop) + (1|indiv2/cross), data = fdata_gather, family = bernoulli,control = list(adapt_delta = 0.95))

summary(baysianGLMERmyc_lesion_GTHR)
bayes_R2(baysianGLMERmyc_lesion_GTHR)

baysianGLMERmyc_lesion_GTHR_1 <- brm(Value ~ Trait * location * species + (1|pop) + (1|indiv2/cross), data = fdata_gather, family = bernoulli,control = list(adapt_delta = 0.95))

baysianGLMERmyc_lesion_GTHR_2 <- brm(Value ~ Trait + (1|pop) + (1|indiv2/cross), data = fdata_gather, family = bernoulli,control = list(adapt_delta = 0.95))

baysianGLMERmyc_lesion_GTHR_NULL <- brm(Value ~ 1 + (1|pop) + (1|indiv2/cross), data = fdata_gather, family = bernoulli,control = list(adapt_delta = 0.95))

model.sel(baysianGLMERmyc_lesion_GTHR,baysianGLMERmyc_lesion_GTHR_1,baysianGLMERmyc_lesion_GTHR_2,baysianGLMERmyc_lesion_GTHR_NULL) #not sure if this function will work with brm model, if not ill just have to do a waic and then a compare like the previous... should test with a smaller model
```
**SUMMARY:**

What is significant with the correlation~location*sp model?

What does it change when we gather the model? What am I really testing?
Does it really test what I want it to test?

What is significant when we gather? Is this model even any good?

###Totals data

```{r Totals model}

# model 
GLMMmyc_lesion <- glmer(binlesion ~ None.Myc * species + (1 | pop) , data = ndatas,family = binomial(link = "logit"))
summary(GLMMmyc_lesion)

options(na.action="na.fail")
dredge(GLMMmyc_lesion)
r.squaredGLMM(GLMMmyc_lesion)
# will need to change this ASAP

ggplot(data = ndata,aes(x = 1-None.Myc, y = 100-None.Path))+
  stat_summary(fun.y=mean, geom="point")+
  geom_smooth(method = "lm")+
  geom_point()+
  theme_simple()+
  facet_wrap("species")+
  labs(x = "Mycorrhizal colonization", y="Lesions")
```
**SUMMARY:**

Is the totals the same as the corr data?

if not what has change? what does it mean biologically: Same/less path=mycorrhiza defend NotSame/less path=mycorrhizal defence is not related to colonization. if theres not an effect then maybe it has more to do with the species composition than the actual mycorrhizal colonization, Thats why we will test pathogen colonization alone (without myc = H2)

## Hypothesis 2: Mycorrhiza & Lesion ~ invasion

### Does location impact the effect of myc on lesion?

```{r location interaction with mycorrhiza and lesion}
GLMMmyc_lesion_loc <- glmer(binlesion ~ None.Myc * location * species + (1 | pop), data = ndatas, family = binomial(link = "logit"))
summary(GLMMmyc_lesion_loc)

emm<-emmeans(GLMMmyc_lesion_loc, list(~None.Myc| species | location, ~None.Myc| location), type = "response" )
summary(emm)

ggplot(data = ndata,aes(x = 1-None.Myc, y = 100-None.Path))+
  geom_smooth(method = "lm")+
  geom_point(aes(color=location))+
  theme_simple()+
  facet_wrap("species")+
  labs(x = "Mycorrhizal colonization", y="Lesions")

pca <- prcomp(ndata[,c(5,9)])
location <- ndata[,13]
library(ggbiplot)
g <- ggbiplot(pca, obs.scale = 1, var.scale = 1, 
                groups = location, ellipse = TRUE, 
                circle = TRUE)
g <- g + scale_color_discrete(name = '')
g <- g + theme(legend.direction = 'horizontal', 
                 legend.position = 'top')
print(g)
```

### Does location impact lesion alone?

```{r location on lesion}
glmerLESION_disp <- glmer (binlesion ~ location * species + (1|pop) + (1|indiv), data = ndatas, family = binomial(link="logit"))
summary(glmerLESION_disp)
options(na.action="na.fail")
dredge(glmerLESION_disp)
r.squaredGLMM(glmerLESION_disp)

ggplot(aes(x = location, y = 100-None.Path, color=location), data=ndata)+
  geom_boxplot()+
  geom_jitter(width=0.1)+
  facet_wrap("species")+
  theme_simple()
```

### Does location impact mycorrhiza alone?

```{r location on mycorrhiza}
glmerMYC_disp <- glmer (binmycorr ~ location * species + (1|pop) + (1|indiv), data = ndata, family = binomial(link="logit"))
summary(glmerMYC_disp)
options(na.action="na.fail")
dredge(glmerMYC_disp)
r.squaredGLMM(glmerMYC_disp)

ggplot(aes(x = location, y = 100-None.Myc, color=location), data=ndata)+
  geom_boxplot()+
  geom_jitter(width=0.1)+
  facet_wrap("species")+
  theme_simple()

```

## Hypothesis 3: Lesion ~ Mycorriza * invasion

```{r Hypothesis 3}

ggplot(data = ndata,aes(x = 1-None.Myc, y = 100-None.Path, color=location))+
  geom_smooth(method = "lm")+
  geom_point()+
  theme_simple()+
  facet_wrap("species")+
  labs(x = "Mycorrhizal colonization", y="Lesions")

ggplot(data = ndata,aes(x = 1-None.Myc, y = 100-None.Path, color=location))+
  geom_smooth(method = "lm")+
  geom_point()+
  theme_simple()+
  labs(x = "Mycorrhizal colonization", y="Lesions")
```

# Molecular data Analysis

## Hypothesis 1: Mycorrhiza_div ~ invasion

## Hypothesis 2: Mycorrhiza_div ~ Pathoge_div