### 2018-03-15 Colonize_Model.Rmd
Draft method for calculating correlation in root cross-sections

---
title: "Root Colonization Model"
author: "Robert I. Colautti"
date: "March 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Colonization of roots by _k_ taxa at each of _N_ root intersections can be modelled as the joint probability of _k_ Bernoulli variables. A simple example is the joint probability of mycorrhizal colonization (Cm) and pathogen colonization (Cp). The joint probabilities are:

$$ P(C_m=n_m, C_p=n_p) $$

$$= P(C_p=n_p | C_m=n_m) \cdot P(C_m=n_m)$$

$$= P(C_m=n_m | C_p=n_p) \cdot P(C_p=n_p)) $$

where n~m~ and n~p~ are 0 or 1.

Now consider the four possible outcomes:

$$A. \space Both := P(C_m = 1 , C_p = 1)$$

$$B. \space Mycorrhizae \space Only := P(C_m = 1 , C_p = 1)$$

$$C. \space Pathogens \space Only := P(C_m = 1 , C_p = 1)$$

$$D. \space Neither := P(C_m = 0, C_p = 0)$$

The variance-covariance matrix for mycorrhizae (m) and pathogens (p) species is:

$$V_m = (A + B)(1 - (A + B))$$

$$V_p = (A + C)(1 - (A + C))$$

$$Cov_{mp} = A - (A + B)(A + C)$$

The correlation coefficient between species is:

$$\frac {Cov_{mp}}{\sqrt{V_m \cdot V_p}} $$

$$= \frac {A - (A + B)(A + C)}{\sqrt {(A + B)(1 - (A + B))(A + C)(1 - (A + C))}}$$