### 2018-03-19 CleanModels.Rmd
Commit work

---
title: "R Notebook"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    highlight: tango
    theme: cerulean
    toc: yes
---

#Housekeeping

##Load packages

```{r Load_packages and functions, message=FALSE, warning=FALSE}
library(tidyverse)
library(lme4)
library(MASS)
library(MuMIn)
library(emmeans)
library(boot)
library(brms)
library(loo)
library(fitdistrplus)
library(mefa)
library(ggbiplot)
library(pvclust)
library(vegan)
library(ggbiplot)
library(factoextra)
library(glmmADMB)
library(car)
library(effects)
library(coefplot2)
library(bbmle)
library(gridExtra)
library(corrplot)
library(ade4)
library(lmtest)
```

##Load costum functions

```{r costum functions}
residfit <- function (model, col="black") {
  f<-fitted(model)
  r<-resid(model, type='pearson')
  plot(f, r, col=col, main='Residuals vs. Fitted')
  L1<-loess(r~f)
  fvec = seq(min(f),max(f),length.out=1000)
  lines(fvec,predict(L1,fvec),col='red')
}

#do the same for the Scale-Location plot
scaleloc <- function(model,col="black") {
  f <- fitted(model)
  r <- sqrt(abs(residuals(model, type='pearson'))) #transformed pearson residuals
  plot(f,r,col=col, main='Scale-Location Plot') 
  L1 <- loess(r~f)
  fvec = seq(min(f),max(f),length.out=1000)
  lines(fvec,predict(L1,fvec),col=2)
}
```

##Load data

### Load/ Modify data: Co-occurence

```{r Introduce co-occurence dataset, message=FALSE, warning=FALSE, echo=FALSE}
mdata <- read.csv("CSV/Assessing_Oct_20.csv",stringsAsFactors=F)
indiv <- mdata$CODE[!mdata$CODE==""]
tdata <- mdata[grep("[0-9]+",mdata$Cross),]
nrow(tdata)/length(indiv)

# This is not the same because some individuals just had the totals not the cooccurence data
## THE COMMENT IS NOT CLEAR, EXPLAIN...
indiv2 <- tdata$CODE[!tdata$CODE==""]
nrow(tdata)/length(indiv2)
indiv2 <- data.frame(indiv2)
indiv2 <- mefa:::rep.data.frame(indiv2,each=100)
fdata <- cbind(indiv2,tdata)
fdata <- fdata[,-c(2,3,4,9,14,17,18)]

# Cleaning up the occurence data
## CLEANING UP HOW?
fdata[is.na(fdata)] <- 0
fdata$Lesion<-rowSums(fdata[,2:4])
fdata[fdata$Lesion>1,]
fdata$Mycorrhiza<-rowSums(fdata[,6:8])
fdata[fdata$Mycorrhiza>1,]
fdata <- fdata[,-c(2:4,6:8)]

#### Checking data for problems ####

summary(fdata)
fdata$None.Path <- as.integer(fdata$None.Path)
fdata$None.Myc <- as.integer(fdata$None.Myc)
fdata[is.na(fdata)] <- 0

sherb <- fdata$Herbivory+fdata$None.Herb
fdata[sherb!=1,]
fdata[sherb!=1,"None.Herb"]<-1

# Are all the sums equal to 100 for herb? 
dframeHerb <- data.frame(counter=NA, Indiv2=NA,Sum=NA)
c <- 1
for (i in unique(fdata$indiv2)){
  dframeHerb[c,] <- NA
  subdata <- fdata[fdata$indiv2==i,]
  dframeHerb$counter[c] <- c
  dframeHerb$Indiv2[c] <- i
  dframeHerb$Sum[c] <- sum(subdata[,"Herbivory"]+subdata[,"None.Herb"])
  subdata <- NA
  c <- c + 1
}
dframeHerb[dframeHerb$Sum!=100,]

# Are all the sums equal to 100 for Mycorrhiza? 
dframeMyc <- data.frame(counter=NA, Indiv2=NA,Sum=NA)
c <- 1
for (i in unique(fdata$indiv2)){
  dframeMyc[c,] <- NA
  subdata <- fdata[fdata$indiv2==i,]
  dframeMyc$counter[c] <- c
  dframeMyc$Indiv2[c] <- i
  dframeMyc$Sum[c] <- sum(subdata[,"Mycorrhiza"]+subdata[,"None.Myc"])
  subdata <- NA
  c <- c + 1
}
dframeMyc[dframeMyc$Sum!=100,]

# Are all the sums equal to 100 for Lesions? 
dframeLesion <- data.frame(counter=NA, Indiv2=NA,Sum=NA)
c <- 1
for (i in unique(fdata$indiv2)){
  dframeLesion[c,] <- NA
  subdata <- fdata[fdata$indiv2==i,]
  dframeLesion$counter[c] <- c
  dframeLesion$Indiv2[c] <- i
  dframeLesion$Sum[c] <- sum(subdata[,"Lesion"]+subdata[,"None.Path"])
  subdata <- NA
  c <- c + 1
}
dframeLesion[dframeLesion$Sum!=100,]

# Separating "Individual" into: pop,location,species columns 
fdata$pop <- gsub("^([0-9]+)[I,O][A-Z]+.*","\\1",fdata$indiv)
fdata$location <- gsub("^[0-9]+([I,O])[A-Z]+.*","\\1",fdata$indiv)
fdata$species <- sub("^[0-9]+[I,O]([A-Z]+).*","\\1",fdata$indiv)
fdata$location <- as.factor(fdata$location)
fdata$species <- as.factor(fdata$species)
fdata$pop <- as.factor(fdata$pop)
fdata$indiv2 <- as.character(fdata$indiv2)
summary(fdata)
fdata %>% drop_na()->fdata

fdata$None.Myc <- as.factor(fdata$None.Myc)
fdata$Herbivory <- as.factor(fdata$Herbivory)
fdata$Mycorrhiza <- as.factor(fdata$Mycorrhiza)
fdata$Lesion <- as.factor(fdata$Lesion)
fdata$None.Herb <- as.factor(fdata$None.Herb)
fdata$None.Path <- as.factor(fdata$None.Path)
summary(fdata)
```

### Load/ Modify data: Totals

```{r Introduce total dataset, message=FALSE, warning=FALSE, echo=FALSE}
mdata<-read.csv("CSV/Assessing_Oct_20.csv",stringsAsFactors=F)
indiv<-mdata$CODE[!mdata$CODE==""]
tdata<-mdata[mdata$Slide=="TOTAL",]
length(indiv)
nrow(tdata)

ndata<-cbind(indiv,tdata)
ndata<-ndata[,-c(2,3,4,9,14,17:37)]
ndata$None.Path<-as.integer(ndata$None.Path)
ndata$Lesions<-as.integer(100-ndata$None.Path)
ndata$None.Myc<-as.integer(ndata$None.Myc)
ndata$Mycorrhiza<-as.integer(100-ndata$None.Myc)

# Checking data for problems
(problemsPath<-ndata[!rowSums(ndata[,c(2:5)])==100,])
(problemsHyph<-ndata[!rowSums(ndata[,c(6:9)])==100,])
(problemsHerb<-ndata[!rowSums(ndata[,c(10:11)])==100,])

# Separating "Individual" into: pop,location,species columns 
ndata$pop<-gsub("^([0-9]+)[I,O][A-Z]+.*","\\1",ndata$indiv)
ndata$location<-gsub("^[0-9]+([I,O])[A-Z]+.*","\\1",ndata$indiv)
ndata$species<-gsub("^[0-9]+[I,O]([A-Z]+).*","\\1",ndata$indiv)
ndata$location<-as.factor(ndata$location)
ndata$species<-as.factor(ndata$species)
ndata$pop<-as.factor(ndata$pop)

#### Totalling the scores####
binlesion<-cbind(ndata$None.Path,100-ndata$None.Path)
binmycorr<-cbind(ndata$None.Myc,100-ndata$None.Myc)
binherbivory<-cbind(ndata$None.Herb,100-ndata$None.Herb)
summary(ndata)
```

### Field survey data

```{r load field data}
FloristicSurvey <-read.csv("CSV/May_2017_FloristicSurvey_summer2016.csv", sep="\t")
PlantLength <- read.csv("CSV/May_2017_PlantLength_summer2016.csv", sep="\t")
Soil_characteristics <- read.csv("CSV/Soil_characteristics.txt", sep="\t")

#Restructuring
head(PlantLength)
str(PlantLength)
colnames(PlantLength)[1] <- "Population"
PlantLength$Population<-as.factor(PlantLength$Population)
PlantLength$Location<-as.factor(PlantLength$Location)

FloristicSurvey$Population<-as.factor(FloristicSurvey$Population) # Make the population a factor and not an interger
FloristicSurvey[is.na(FloristicSurvey)] <- 0
sum(row.has.na <- apply(FloristicSurvey, 1, function(x){any(is.na(x))}))
```

### Root size

```{r load root size data}
Root_size <-read.csv("CSV/Root_size.csv")
Root_size$pop<-gsub("^([0-9]+)[I,O][A-Z]+.*","\\1",Root_size$indiv)
Root_size$location<-gsub("^[0-9]+([I,O])[A-Z]+.*","\\1",Root_size$indiv)
Root_size$species<-gsub("^[0-9]+[I,O]([A-Z]+).*","\\1",Root_size$indiv)
Root_size$location<-as.factor(Root_size$location)
Root_size$species<-as.factor(Root_size$species)
Root_size$pop<-as.factor(Root_size$pop)
```

###Make a full dataset

```{r full data set}
Soil_characteristics$Location<-as.character(Soil_characteristics$Location)
for(i in 1:length(Soil_characteristics$Location))if(Soil_characteristics[i,2]=="in"){
   Soil_characteristics[i,2]<-"I"
}else{
  Soil_characteristics[i,2]<-"O"
}
Soil_characteristics$Location<-as.factor(Soil_characteristics$Location)  
names(Soil_characteristics)[1:2]<-c("pop","location")

fulldata <- merge(Soil_characteristics, ndata, by=c("pop","location"))
names(Root_size)[1]<-"indiv"
fulldata <- merge(fulldata, Root_size, by=c("indiv","location","pop","species"))
for(i in 1:length(fulldata$indiv)){
fulldata$mean_rootsize[i]<-(sum(fulldata[i,20:29])/length(fulldata[,20:29]))
}
fulldata<-fulldata[,-c(20:29)]

binlesionF<-cbind(fulldata$None.Path,100-fulldata$None.Path)
binmycorrF<-cbind(fulldata$None.Myc,100-fulldata$None.Myc)
binherbivoryF<-cbind(fulldata$None.Herb,100-fulldata$None.Herb)

fulldata$pop<-as.factor(fulldata$pop)
fulldata$Mycorrhiza<-100-fulldata$None.Myc
fulldata$Lesion<-100-fulldata$None.Path
```

#Variable descriptions

###Co-occurence

  * **indiv2:** The unique code given to a sample. In this dataset the unique sample code repeats 100 times for each cross where an observation was taken on the root sample.

  * **None.Path:** A binary representation of whether a sign of pathogen activity was recorded (0) or not (1).

  * **Lesion:** The counterpart to the previous variable (None.Path). A binary representation of whether a sign of pathogen activity was recorded (1) or not (0).

  * **None.Myc:** A binary representation of whether a sign of myccorhizal activity was recorded (0) or not (1).

  * **Mycorrhiza:** The counterpart to the previous variable (None.Myc). A binary representation of whether a sign of mycorrhizal activity was recorded (1) or not (0).

  * **None.Herb:** A binary representation of whether herbivory was recorded (0) or not (1).

  * **Herbivory:** The counterpart to the previous variable (None.Herb). A binary representation of whether Herbivoryn was recorded (1) or not (0).

  * **Population (pop):** The coding number representing the population at which the sample was collected.

  * **location:** A code representing the whether the sample was collected inside a Alliaria petiolata population (I) or whther it was collected at least 7 m outside of he furthest individual in the A. petiolata population (O).

  * **species:** The particular species to which the sample belongs.

  * **Cross:** The particular cross number where the data was recorded on the individual sample.

###Total scores

  * **indiv:** The unique code given to a sample.

  * **Decay:** The total number of crosses where decay was recorded an individual sample when doing a total of 100 crosses/ sample.

  * **Pathogen:** The total number of crosses where pathogen was recorded on an individual sample when doing a total of 100 crosses/ sample.

  * **Hyphae:** The total number of crosses where non-myccorhizal hyphae was recorded on an individual sample when doing a total of 100 crosses/ sample.

  * **None.Path:** The total number of crosses where no signs of pathogen activities were recorded on an individual sample when doing a total of 100 crosses/ sample. Note that the total of Decay, Pathogen, Hyphae, and None.Path must come to 100 per individual to account for all 100 crosses.

  * **Arbuscule:** The total number of crosses where an arbuscule was recorded on an individual sample when doing a total of 100 crosses/ sample.

  * **Vesicules:** The total number of crosses where a vesicule was recorded on an individual sample when doing a total of 100 crosses/ sample.

  * **M_Hyphae:** The total number of crosses where myccorhizal hyphae was recorded on an individual sample when doing a total of 100 crosses/ sample.

  * **None.Myc:** The total number of crosses where no signs of mycorrhizal activities were recorded on an individual sample when doing a total of 100 crosses/ sample. Note that the total of Arbuscules, Vesicules, M_Hyphae, and None.Myc must come to 100 per individual to account for all 100 crosses.

  * **Herbivory:** The total number of crosses where herbivory was recorded on an individual sample when doing a total of 100 crosses/ sample.

  * **None.Herb:** The total number of crosses where no signs of herbivory was recorded on an individual sample when doing a total of 100 crosses/ sample. Note that the total of Herbivory, and None.Herb must come to 100 per individual to account for all 100 crosses.

  * **Population (pop):** The coding number representing the population at which the sample was collected.

  * **location:** A code representing the whether the sample was collected inside a Alliaria petiolata population (I) or whther it was collected at least 7 m outside of he furthest individual in the A. petiolata population (O).

  * **species:** The particular species to which the sample belongs.

###Floristic Survey

###Soil Characteristics

###Root size

#Soil abiotic qualities and location

### Model selection

```{r model fittin soil}
var<-Soil_characteristics[,c(1,2)]
Y<-Soil_characteristics[,-c(1,2,5)]
var$pop<-as.factor(var$pop)
Location<-as.factor(var$location)
Y$Sample <-paste(var$pop,var$location) 
rownames(Y)<-Y$Sample
Y$Sample<-NULL

covar_cor = cor(Y)
corrplot.mixed(covar_cor, upper = "ellipse", lower = "number")
Nutrient_PCA = dudi.pca(df = Y[, c(3, 4)], scannf = FALSE, nf = 1)
Nutrient_PCA$co
Nutrient_PCA$eig/sum(Nutrient_PCA$eig)
Soil_characteristics$Nutrient_PC1 = Nutrient_PCA$li[,1]

dat<-data.frame(c(1:20))
dat[1]<-Soil_characteristics$location
colnames(dat)[1]<-"Location"
dat$Agg_stability.s<-scale(Soil_characteristics$Agg_stability)
dat$pH.s<-scale(Soil_characteristics$pH)
dat$Nutrient_PC1.s<-scale(Soil_characteristics$Nutrient_PC1)

GLM_soil <- glm(Location~Agg_stability.s*pH.s*Nutrient_PC1.s,
                    family=binomial(link='logit'),
                    data=dat)
```

###LRT

```{r LRT soil}
GLM_soil2 <- glm(Location~Agg_stability.s+pH.s+Nutrient_PC1.s+
                   Agg_stability.s:pH.s + Agg_stability.s:Nutrient_PC1.s+
                   Nutrient_PC1.s:pH.s,
                family=binomial(link='logit'),
                data=dat)
anova(GLM_soil,GLM_soil2)

GLM_soil3 <- glm(Location~Agg_stability.s+pH.s+Nutrient_PC1.s+
                   Agg_stability.s:pH.s + Agg_stability.s:Nutrient_PC1.s,
                 family=binomial(link='logit'),
                 data=dat)
anova(GLM_soil2,GLM_soil3)

GLM_soil4 <- glm(Location~Agg_stability.s+pH.s+Nutrient_PC1.s+
                   Agg_stability.s:pH.s,
                 family=binomial(link='logit'),
                 data=dat)
anova(GLM_soil3,GLM_soil4)

GLM_soil5 <- glm(Location~Agg_stability.s+pH.s+Nutrient_PC1.s,
                 family=binomial(link='logit'),
                 data=dat)
anova(GLM_soil4,GLM_soil5)

GLM_soil6 <- glm(Location~Agg_stability.s+pH.s,
                 family=binomial(link='logit'),
                 data=dat)
anova(GLM_soil5,GLM_soil6)

GLM_soil7 <- glm(Location~Agg_stability.s,
                 family=binomial(link='logit'),
                 data=dat)
anova(GLM_soil6,GLM_soil7)

GLM_soil8 <- glm(Location~1,
                 family=binomial(link='logit'),
                 data=dat)
anova(GLM_soil7,GLM_soil8)
```

###descriptive stats

```{r descriptive stats soil}
summary(GLM_soil)
confint(GLM_soil)

pr<-resid(GLM_soil, type='pearson')
n<-nrow(Soil_characteristics)
c<-length(coef(GLM_soil))
(Disp<-sum(pr^2)/(n-c))
```

#Garlic mustard on Mycorrhiza

###Model selection

```{r Model fit mycGM gauss}
hist(ndata$Lesions)
# In vs Out
ggplot(data=ndata,aes(x=location))+geom_histogram(stat="count")+facet_wrap(~pop)
ggplot(data=ndata,aes(x=location))+geom_histogram(stat="count")+facet_wrap(~species)
# Each Species
ggplot(data=ndata,aes(x=species))+geom_histogram(stat="count")+facet_wrap(~pop)
# Each Population
ggplot(data=ndata,aes(x=pop))+geom_histogram(stat="count")

ndata$Prop_Lesions<- ndata$Lesions/(ndata$Lesions+ndata$None.Path)
hist(ndata$Prop_Lesions)
shapiro.test(log(ndata$Prop_Lesions))
mod1.gauss<-lmer(Prop_Lesions~ species + location + species:location +
                       (1|pop),
                     data=ndata) 
mod1.gauss<-lm(Prop_Lesions~ species + location + species:location,data=ndata)
bc<-boxcox(mod1.gauss, lambda = seq(-2, 2, 0.1))
lambda <- bc$x[which.max(bc$y)]
ndata <- cbind(ndata, ((ndata$Prop_Lesions^lambda)-1)/lambda)
names(ndata)[length(ndata)] <- "Yprime"
shapiro.test(ndata$Yprime)
hist(ndata$Yprime)
scaleloc(mod1.gauss)
residfit(mod1.gauss)

mod1.gaussY<-glmmadmb(Yprime~ species + location + species:location +
                       (1|pop),
                     data=ndata, family="gaussian")
qqPlot(ndata$Yprime)

cooksd <- cooks.distance(mod1.gauss)
names(cooksd) <- 1:length(cooksd)
plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 4*mean(cooksd, na.rm=T), col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4*mean(cooksd, na.rm=T),names(cooksd),""), col="red")

ndataRM<-ndata[-c(66,73,97,130,133,135,136,176,172,185,204),]
qqPlot(ndataRM$Yprime)
shapiro.test(ndataRM$Yprime)

mod1.gaussYRM<-glmmadmb(Yprime~ species + location + species:location +
                       (1|pop),
                     data=ndataRM, family="gaussian")

bptest(mod1.gaussYRM)
hist(ndataRM$Yprime)
mod1.gaussYRM<-lmer(Yprime~ species + location + species:location +
                       (1|pop),
                     data=ndataRM)
scaleloc(mod1.gaussYRM)
residfit(mod1.gaussYRM)

ICtab(mod1.gaussY,mod1.poisson,mod1.nb,mod1.gaussYRM)
```
```{r Model fit mycGM beta}
descdist(ndata$Prop_Lesions,boot=1000)
summary(ndata$Prop_Lesions)
plot(density(ndata$Prop_Lesions))
mu <- mean(ndata$Prop_Lesions)
var <- var(ndata$Prop_Lesions)
estBetaParams <- function(mu, var) {
  alpha <- ((1 - mu) / var - 1 / mu) * mu ^ 2
  beta <- alpha * (1 / mu - 1)
  return(params = list(alpha = alpha, beta = beta))
}
estBetaParams(mu, var)

mod1.beta<-glmmTMB(Prop_Lesions~ species + location + species:location +
                       (1|pop), ziformula=,
                     data=ndata, family=list(family="beta",link="logit"))

```

```{r Model fit mycGM binom}
mod1.binom<-glmmadmb(binmycorr ~ species + location + species:location +
                       (1|pop),
                     data=ndata, family="binomial")

pr<-resid(mod1.binom, type='pearson')
(phimod1.binom <- sum(pr^2)/df.residual(mod1.binom))

scaleloc(mod1.binom)
residfit(mod1.binom)

```

```{r Model fit mycGM poisson}
mod1.poisson <-glmmadmb(Mycorrhiza ~ species + location + species:location +
                         (1|pop),
                       data=ndata, family="poisson")

pr<-resid(mod1.poisson, type='pearson')
(phi <- sum(pr^2)/df.residual(mod1.poisson))

scaleloc(mod1.poisson)
residfit(mod1.poisson)

ChiCrit<-qchisq(0.95, df.residual(mod1.poisson))
deviance(mod1.poisson)
ChiSq<-sum(pr^2)

ChiCrit>=ChiSq
```

```{r Model fit mycGM negBinom}
# http://data.princeton.edu/wws509/r/overdispersion.html

mod1.nb<-glmmadmb(Mycorrhiza ~ species + location + species:location +
                    (1|pop),
                  data=ndata,family="nbinom1")

-2*(logLik(mod1.poisson)-logLik(mod1.nb))

scaleloc(mod1.nb)
residfit(mod1.nb)

pr<-resid(mod1.nb, type='pearson')
ChiSq<-sum(pr^2)
ChiCrit>=ChiSq #still too big but much smaller...

mod1.nb1<-glmer.nb(Mycorrhiza ~ species + location + species:location +
                    (1|pop),
                  data=ndata, control=glmerControl(optimizer="bobyqa")) 
theta<-getME(mod1.nb1, "glmer.nb.theta")

xb <- predict(mod1.poisson)
g <- cut(xb, breaks=quantile(xb,seq(0,100,5)/100))
m <- tapply(ndata$Mycorrhiza, g, mean)
v <- tapply(ndata$Mycorrhiza, g, var)

plot(m, v, xlab="Mean", ylab="Variance",main="Mean-Variance Relationship")
x <- seq(42.16667,83.77778,0.3)
lines(x, x*phi, lty="dashed")
lines(x, x*(1+x/theta))

```

###LRT

```{r LRT mycGM}
mod1.1<-glmmadmb(Mycorrhiza ~ species + location + 
                    (1|pop),
                  data=ndata, family="nbinom1") 
# species:location?
anova(mod1.1, mod1.nb) #discard

mod1.2<-glmmadmb(Mycorrhiza ~ species + 
                    (1|pop),
                  data=ndata, family="nbinom1") 
# location?
anova(mod1.2, mod1.1) #discard????

mod1.3<-glmmadmb(Mycorrhiza ~ location + 
                    (1|pop),
                  data=ndata, family="nbinom1")

mod1.4<-glmmadmb(Mycorrhiza ~ 1 + 
                    (1|pop),
                  data=ndata, family="nbinom1")

anova(mod1.4 , mod1.3)
anova(mod1.4 , mod1.2)

options(na.action = "na.fail")
dredge(mod1.nb)
```

###Descriptive stats

```{r descriptive stats mycGM}
pr<-resid(mod1.1, type='pearson')
n<-nrow(ndata)
c<-length(coef(mod1.1))
(Disp<-sum(pr^2)/(n-c))

summary(mod1.1)
confint(mod1.1)

# no efffect on colonization... maybe the plantswere there before GM?
coefplot2(mod1.1)
```

#Garlic mustard on Lesion

###Model selection

```{r model sel lesionGM}
hist(ndata$Lesions)

mod2.binom<-glmmadmb(binlesion ~ species + location + species:location +
                       (1|pop),
                     data=ndata, family="binomial")

pr<-resid(mod2.binom, type='pearson')
n<-nrow(ndata)
c<-length(coef(mod2.binom))
(Dispmod2.binom<-sum(pr^2)/(n-c))
scaleloc(mod2.binom)
residfit(mod2.binom)

mod2.binom.ind<-glmmadmb(binlesion ~ species + location + species:location +
                       (1|pop)+ (1|indiv),
                     data=ndata, family="binomial")

pr<-resid(mod2.binom.ind, type='pearson')
n<-nrow(ndata)
c<-length(coef(mod2.binom.ind))
(Dispmod2.binom.ind<-sum(pr^2)/(n-c))
scaleloc(mod2.binom.ind)
residfit(mod2.binom.ind)

mod2.binom.sp<-glmmadmb(binlesion ~ species + location + species:location +
                 (1|pop) + (species|pop),
               data=ndata, family="binomial")

pr<-resid(mod2.binom.sp, type='pearson')
n<-nrow(ndata)
c<-length(coef(mod2.binom.sp))
(Disp.sp<-sum(pr^2)/(n-c))
scaleloc(mod2.binom.sp)
residfit(mod2.binom.sp)

mod2.poisson<-glmmadmb(Lesions ~ species + location + species:location +
                    (1|pop),
                  data=ndata, family="poisson")

pr<-resid(mod2.poisson, type='pearson')
n<-nrow(ndata)
c<-length(coef(mod2.poisson))
(Dispmod2.poisson<-sum(pr^2)/(n-c))
scaleloc(mod2.poisson)
residfit(mod2.poisson)

mod2.nb.sp<-glmmadmb(Lesions ~ species + location + species:location +
                  (species|pop) + (1|pop),
                  data=ndata, family="nbinom1")

pr<-resid(mod2.nb.sp, type='pearson')
n<-nrow(ndata)
c<-length(coef(mod2.nb.sp))
(Dispmod1.nb.sp<-sum(pr^2)/(n-c))
scaleloc(mod2.nb.sp)
residfit(mod2.nb.sp)

mod2.nb<-glmmadmb(Lesions ~ species + location + species:location +
                    (1|pop),
                  data=ndata, family="nbinom1")

pr<-resid(mod2.nb, type='pearson')
n<-nrow(ndata)
c<-length(coef(mod2.nb))
(Dispmod2.nb<-sum(pr^2)/(n-c))
scaleloc(mod2.nb)
residfit(mod2.nb)

ICtab(mod2.poisson,mod2.binom,mod2.binom.sp,mod2.nb.sp,mod2.nb)
```

###LRT

```{r LRT lesionGM}
mod2.1<-glmmadmb(Lesions ~ species + location +
                    (1|pop),
                  data=ndata, family="nbinom1") 
anova(mod2.1, mod2.nb)

mod2.2<-glmmadmb(Lesions ~ species +
                    (1|pop),
                  data=ndata, family="nbinom1")

mod2.3<-glmmadmb(Lesions ~ location +
                    (1|pop),
                  data=ndata, family="nbinom1")

mod2.4<-glmmadmb(Lesions ~ 1 +
                    (1|pop),
                  data=ndata, family="nbinom1") 
anova(mod2.4 , mod2.2)
anova(mod2.4 , mod2.3)

dredge( mod2.nb)
```

###Descriptive stats

```{r descriptive stats lesionGM}
pr<-resid(mod2.nb, type='pearson')
n<-nrow(ndata)
c<-length(coef(mod2.nb))
(Disp<-sum(pr^2)/(n-c))

summary(mod2.nb)
confint(mod2.nb)

coefplot2(mod2.nb)
```

# Myc and lesion Correlation Model

*NOTE: Yprime = corr coefficient of Myc and Lesion at every cross per indiv*

###Model selection

```{r}
# calculating the r
Root_size$mean_rootsize <- rowSums(Root_size[,2:11])/length(Root_size[,2:11])
names(Root_size)[1]<-"indiv2"

fdata_rs <- merge(fdata, Root_size, by=c("indiv2","location","pop","species"))
indivCODE <- as.character(unique(fdata_rs$indiv2))
dfcorr <-data.frame(indiv2=c(0),r=c(0))
for (i in 1:length(indivCODE)){
  dat1<-NULL
  dat1<-fdata_rs[fdata_rs$indiv2==indivCODE[i],]
  dfcorr[i,1] <- indivCODE[i]
  dfcorr[i,2] <- cor(as.numeric(dat1$Mycorrhiza),as.numeric(dat1$Lesion))
}
dfcorr %>% drop_na()->dfcorr

fdata_rs.corr <- merge(dfcorr, Root_size, by="indiv2")
shapiro.test(fdata_rs.corr$r)

fdata_rs.corr$r.t<-fdata_rs.corr$r+1
CorrLM<-lm(r.t~ species * location * mean_rootsize,data=fdata_rs.corr)

bc<-boxcox(CorrLM, lambda = seq(-2, 2, 0.1))
lambda <- bc$x[which.max(bc$y)]
fdata_rs.corr <- cbind(fdata_rs.corr, ((fdata_rs.corr$r.t^lambda)-1)/lambda)
names(fdata_rs.corr)[length(fdata_rs.corr)] <- "Yprime"
shapiro.test(fdata_rs.corr$Yprime)
hist(fdata_rs.corr$Yprime)

CorrLMERBC<-lmer(Yprime~ species + location + mean_rootsize + 
                   location:species +
                   (1|pop),
                 data=fdata_rs.corr)
pr<-resid(CorrLMERBC, type='pearson')
n<-nrow(fdata_rs.corr)
c<-length(coef(CorrLMERBC))
(Disp<-sum(pr^2)/(n-c))

plot(effect("location:mean_rootsize", CorrLMERBC))
plot(effect("mean_rootsize", CorrLMERBC))
```

####LRT

```{r LRT corrmod}
CorrLMERBC<-lmer(Yprime~ species + location + mean_rootsize + 
                   location:species +
                   (1|pop),
                 data=fdata_rs.corr)

CorrLMERBC1<-lmer(Yprime~ species + location + mean_rootsize+
                   (1|pop),
                 data=fdata_rs.corr)
anova(CorrLMERBC1, CorrLMERBC)

CorrLMERBC2<-lmer(Yprime~ species + location + 
                   (1|pop),
                 data=fdata_rs.corr)
anova(CorrLMERBC2, CorrLMERBC1)

CorrLMERBC3<-lmer(Yprime~ species +  
                    (1|pop),
                  data=fdata_rs.corr)
anova(CorrLMERBC3, CorrLMERBC2)

CorrLMERBC4<-lmer(Yprime~ 1+ 
                    (1|pop),
                  data=fdata_rs.corr)
anova(CorrLMERBC3, CorrLMERBC4)
```

###Descriptive stats

```{r descriptive stats cormod}
pr<-resid(CorrLMERBC, type='pearson')
n<-nrow(fdata_rs.corr)
c<-length(coef(CorrLMERBC))
(Disp<-sum(pr^2)/(n-c))

summary(CorrLMERBC)
confint(CorrLMERBC)
```

#Everything on lesion

###Model selection

```{r model selection fullmod}
var<-fulldata[,c(5,6,8,9,22,23,24)]
covar_cor = cor(var)
corrplot.mixed(covar_cor, upper = "ellipse", lower = "number")
Nutrient_PCA = dudi.pca(df = var[, c(3, 4)], scannf = FALSE, nf = 1)
Nutrient_PCA$co
Nutrient_PCA$eig/sum(Nutrient_PCA$eig)
fulldata$Nutrient_PC1 = Nutrient_PCA$li[,1]

mod10.nbinom <- glmmadmb(Lesion ~ Mycorrhiza + scale(mean_rootsize) + species + 
                      scale(pH)+ scale(Agg_stability) + Nutrient_PC1 + 
                      (1 | pop)+
                      Mycorrhiza:Nutrient_PC1 + Mycorrhiza:scale(mean_rootsize) +
                      Mycorrhiza:species + Mycorrhiza:scale(pH) +
                      Mycorrhiza:scale(Agg_stability), 
                    data = fulldata,family = "nbinom1")
mod10.nbinom.ind <- glmmadmb(Lesion ~ Mycorrhiza + scale(mean_rootsize) + species + 
                      scale(pH)+ scale(Agg_stability) + Nutrient_PC1 + 
                      (1 | pop)+(1 | indiv)+
                      Mycorrhiza:Nutrient_PC1 + Mycorrhiza:scale(mean_rootsize) +
                      Mycorrhiza:species + Mycorrhiza:scale(pH) +
                      Mycorrhiza:scale(Agg_stability), 
                    data = fulldata,family = "nbinom1")

# mod10.poisson <- glmmadmb(Lesion ~ Mycorrhiza + scale(mean_rootsize) + species + 
                      # scale(pH)+ scale(Agg_stability) + Nutrient_PC1 + 
                      # (1 | pop)+
                      # Mycorrhiza:Nutrient_PC1 + Mycorrhiza:scale(mean_rootsize) +
                      # Mycorrhiza:species + Mycorrhiza:scale(pH) +
                      # Mycorrhiza:scale(Agg_stability), 
                    # data = fulldata,family = "poisson") 
#fails
mod10.poisson.ind <- glmmadmb(Lesion ~ Mycorrhiza + scale(mean_rootsize) + species + 
                      scale(pH)+ scale(Agg_stability) + Nutrient_PC1 + 
                      (1 | pop)+(1 | indiv)+
                      Mycorrhiza:Nutrient_PC1 + Mycorrhiza:scale(mean_rootsize) +
                      Mycorrhiza:species + Mycorrhiza:scale(pH) +
                      Mycorrhiza:scale(Agg_stability), 
                    data = fulldata,family = "poisson")

mod10.binom <- glmmadmb(binlesionF ~ Mycorrhiza + scale(mean_rootsize) + species + 
                      scale(pH)+ scale(Agg_stability) + Nutrient_PC1 + 
                      (1 | pop)+
                      Mycorrhiza:Nutrient_PC1 + Mycorrhiza:scale(mean_rootsize) +
                      Mycorrhiza:species + Mycorrhiza:scale(pH) +
                      Mycorrhiza:scale(Agg_stability), 
                    data = fulldata,family = "binomial")
# mod10.binom.ind <- glmmadmb(binlesionF ~ Mycorrhiza + scale(mean_rootsize) + species + 
                      # scale(pH)+ scale(Agg_stability) + Nutrient_PC1 + 
                      # (1 | pop) + (1 | indiv)+
                      # Mycorrhiza:Nutrient_PC1 + Mycorrhiza:scale(mean_rootsize) +
                      # Mycorrhiza:species + Mycorrhiza:scale(pH) +
                      # Mycorrhiza:scale(Agg_stability), 
                    # data = fulldata,family = "binomial") 
#fail

ICtab(mod10.binom, mod10.poisson.ind, mod10.nbinom, mod10.nbinom.ind)
```

###LRT

*cant include location because when pop is a random factor locatin is correlated with all the other measurements! They were taken "in" and "out".*

```{r LRT fullmodel}
mod10.binom <- glmmadmb(binlesionF ~ Mycorrhiza + scale(mean_rootsize) + species + 
                      scale(pH)+ scale(Agg_stability) + Nutrient_PC1 + 
                      (1 | pop)+
                      Mycorrhiza:Nutrient_PC1 + Mycorrhiza:scale(mean_rootsize) +
                      Mycorrhiza:species + Mycorrhiza:scale(pH) +
                      Mycorrhiza:scale(Agg_stability), 
                    data = fulldata,family = "binomial")

mod10.1 <- glmmadmb(binlesionF ~ Mycorrhiza + scale(mean_rootsize) + species + 
                      scale(pH)+ scale(Agg_stability) + Nutrient_PC1 + 
                      (1 | pop)+
                      Mycorrhiza:Nutrient_PC1 + Mycorrhiza:scale(mean_rootsize) +
                      Mycorrhiza:species + Mycorrhiza:scale(pH), 
                    data = fulldata,family = "binomial")
# Mycorrhiza:scale(Agg_stability)? 
anova(mod10.1, mod10.binom) # important

mod10.2 <- glmmadmb(binlesionF ~ Mycorrhiza + scale(mean_rootsize) + species + 
                      scale(pH)+ scale(Agg_stability) + Nutrient_PC1 + 
                      (1 | pop)+
                      Mycorrhiza:Nutrient_PC1 + Mycorrhiza:scale(mean_rootsize) +
                      Mycorrhiza:species , 
                    data = fulldata,family = "binomial")
# Mycorrhiza:scale(pH)? 
anova(mod10.2, mod10.1) # important

mod10.3 <- glmmadmb(binlesionF ~ Mycorrhiza + scale(mean_rootsize) + species + 
                      scale(pH)+ scale(Agg_stability) + Nutrient_PC1 + 
                      (1 | pop)+
                      Mycorrhiza:Nutrient_PC1 + Mycorrhiza:scale(mean_rootsize) +
                      Mycorrhiza:scale(pH), 
                    data = fulldata,family = "binomial")
# Mycorrhiza:species? 
anova(mod10.3, mod10.1) # important

mod10.4 <- glmmadmb(binlesionF ~ Mycorrhiza + scale(mean_rootsize) + species + 
                      scale(pH)+ scale(Agg_stability) + Nutrient_PC1 + 
                      (1 | pop)+
                      Mycorrhiza:Nutrient_PC1+
                      Mycorrhiza:species + Mycorrhiza:scale(pH), 
                    data = fulldata,family = "binomial")
# Mycorrhiza:scale(mean_rootsize)?
anova(mod10.4,mod10.1) #keep

mod10.1 <- glmmadmb(binlesionF ~ Mycorrhiza + scale(mean_rootsize) + species + 
                      scale(pH)+ scale(Agg_stability) + Nutrient_PC1 + 
                      (1 | pop)+
                      Mycorrhiza:scale(mean_rootsize) +
                      Mycorrhiza:species + Mycorrhiza:scale(pH), 
                    data = fulldata,family = "binomial")
# Mycorrhiza:Nutrient_PC1?
anova(mod10.3,mod10.4) #keep nutrient myc interaction

mod10.5 <- glmmadmb(binlesionF ~ Mycorrhiza + scale(mean_rootsize) + species + 
                      scale(pH)+ scale(Agg_stability) + 
                      (1 | pop)+
                      Mycorrhiza:Nutrient_PC1 + Mycorrhiza:scale(mean_rootsize) +
                      Mycorrhiza:species + Mycorrhiza:scale(pH) +
                      Mycorrhiza:scale(Agg_stability), 
                    data = fulldata,family = "binomial")
# Nutrient_PC1?
anova(mod10.5, mod10.binom) #important

mod10.6 <- glmmadmb(binlesionF ~ Mycorrhiza + scale(mean_rootsize) + species + 
                      scale(pH) + Nutrient_PC1 + 
                      (1 | pop)+
                      Mycorrhiza:Nutrient_PC1 + Mycorrhiza:scale(mean_rootsize) +
                      Mycorrhiza:species + Mycorrhiza:scale(pH) +
                      Mycorrhiza:scale(Agg_stability), 
                    data = fulldata,family = "binomial")
# scale(Agg_stability)?
anova(mod10.6 ,mod10.binom) #NOPE

mod10.7 <- glmmadmb(binlesionF ~ Mycorrhiza + scale(mean_rootsize) + species + 
                      Nutrient_PC1 + 
                      (1 | pop)+
                      Mycorrhiza:Nutrient_PC1 + Mycorrhiza:scale(mean_rootsize) +
                      Mycorrhiza:species + Mycorrhiza:scale(pH) +
                      Mycorrhiza:scale(Agg_stability), 
                    data = fulldata,family = "binomial")
# scale(pH)?
anova(mod10.7,mod10.6) #NOPE

mod10.8 <- glmmadmb(binlesionF ~ Mycorrhiza + scale(mean_rootsize) +
                      Nutrient_PC1 + 
                      (1 | pop)+
                      Mycorrhiza:Nutrient_PC1 + Mycorrhiza:scale(mean_rootsize) +
                      Mycorrhiza:species + Mycorrhiza:scale(pH) +
                      Mycorrhiza:scale(Agg_stability), 
                    data = fulldata,family = "binomial")
# Species?
anova(mod10.7,mod10.8) #keep

mod10.9 <- glmmadmb(binlesionF ~ Mycorrhiza + species + 
                      Nutrient_PC1 + 
                      (1 | pop)+
                      Mycorrhiza:Nutrient_PC1 + Mycorrhiza:scale(mean_rootsize) +
                      Mycorrhiza:species + Mycorrhiza:scale(pH) +
                      Mycorrhiza:scale(Agg_stability), 
                    data = fulldata,family = "binomial")
# scale(mean_rootsize)?
anova(mod10.9,mod10.7) #Keep

mod10.10 <- glmmadmb(binlesionF ~ scale(mean_rootsize) + species + 
                      Nutrient_PC1 + 
                      (1 | pop)+
                      Mycorrhiza:Nutrient_PC1 + Mycorrhiza:scale(mean_rootsize) +
                      Mycorrhiza:species + Mycorrhiza:scale(pH) +
                      Mycorrhiza:scale(Agg_stability), 
                    data = fulldata,family = "binomial")

# Mycorrhiza?
anova(mod10.10,mod10.7) #keep
```

###Descriptive stats

```{r desctiptive stats fullmod}
pr<-resid(mod10.3, type='pearson')
n<-nrow(fulldata)
c<-length(coef(mod10.3))
(Disp<-sum(pr^2)/(n-c))

summary(mod10.7)
confint(mod10.7)
coefplot2(mod10.7)
```